\chapter*{Acknowledgement}
\markboth{Acknowledgement}{Acknowledgement}

{\singlespace

First and foremost, I would like to think Frank Doyle III for taking the time to advise me and for providing numerous opportunities for me to present my research at conferences.
I also thank the other members of my PhD thesis committee for their continued support and encouragement: Samir Mitragotri and Michelle O'Malley for keeping my work biologically relevant; and Linda Petzold, for her expertise in numerical methods.
I would also like to thank Linda for her excellent class in ODE integration, which greatly helped me understand and appreciate the algorithms I had largely taken for granted.

I would also like to thank the past and current members of the Doyle group, who have made my past five years so memorable.
In no particular order, I thank Stephanie Taylor, for while we didn't actually overlap at UCSB, she nevertheless provided constant guidance and suggestions on the various manuscripts and presentations I prepared during my PhD thesis.
I also thank Bharath Ananthasubramaniam and Panagiota (Pegy) Foteinou for their help getting me started when I first joined the group - lunches haven't been the same since you guys left.
I also thank John Abel for his overall enthusiasm and specifically his help with stochastic methods, without which I may have taken a different path for my last two papers.
I should not forget group members who have worked on other projects: Ravi Gondhalekar, Yongqiang Wang, and Felipe N\'{u}\~{n}ez, who graciously put up with my constant questions about spectral analysis and analytical methods.

I would also like to thank my fianc\'{e}e Katherine LeVan - first in a professional context, for her depth of knowledge of classical statistics and assistance with many aspects of my research.
Most importantly, of course, I thank her for her endless patience and support. 
Similarly, I thank my parents and brother, Christopher, for always being there when I need them.

Finally, I thank the various funding sources who have made this project possible.
Most importantly, I appreciate the generous fellowship given to me by the Mitsubishi Chemical Corporation, which has given me the freedom to pursue whichever project caught my enthusiasm.
I also thank financial support from the National Institutes of Health/National Institute of General Medical Sciences under award number 1R01GM096873-01 and from the Institute for Collaborative Biotechnologies through grant W911NF-09-0001 from the U.S.\ Army Research Office.

}

%Previous group members - Bharath, Stephanie, Pegy, John
%Family
%Money (include Mitsubishi)

% \Blindtext
